---
title: "Run models for Poulin et al. 'Sociality helps navigate anthropogenic risks: evidence from elk crossing a major highway'"
author: "Marie-Pier Poulin"
date: "2023-10-02"
output: html_document
---

# Description of the data, code and file structure

The following code allows to run models to investigate the effect of sociality on the probability of crossing a major highway by elk in Yoho National Park (British Columbia). The dataset is described below, with details on column names and units of each variable.

Data: PoulinEtAl2023-hourlyTravelSteps_wSocialPhenotypes.csv

This dataset includes all observed hourly steps classified as 'traveling' by the HMM, during which at least 2 collared elk were together (i.e., within 100-m of each other during simultaneous fixes within a 15-min interval). It also includes characteristics such as traffic volume recorded on the Trans-Canada highway during the step, season, and social phenotypes of the focal individual and of its group. Below is a description of the columns in the .csv file:

- elkID: Tag number of the focal individual (i.e., elk that performed the movement step).
- winterSeason: Extended winter season (i.e., period when elk were on their winter range or range near the road) // w2 = 2019-2020 and w3 = 2020-2021.
- elkID_winterSeason: Unique combination of the focal individual and extended winter season - used as a random intercept in our models to account for pseudoreplication within individuals and winter seasons.
- season: Identifies if the step was during Fall (August to November), Winter (December to February), or Spring (March to June).

- hwyCros: Highway crossing // 0 = the step did not cross the Trans-Canada highway (TCH) and 1 = the step crossed the TCH.
- trafficVol: Hourly traffic volume on the TCH for the day and hour at which the step was observed.
- trafficVol100: Hourly traffic volume (in hundreds of cars per hour) for the day and hour at which the step was observed.

- nbrCollars: Number of collared elk in the group.
- propCollars: Proportion of all collars that were deployed during the day of the steps that were in the group.
- groupSize: Predicted total group size, based on models of total group size from propCollars using field observations of groups (see Appendix S2 for more info).

[Social phenotypes of individuals]

- indivDominance: Mean elo-score of the focal individual, based on observations of agonistic interactions between pairs of collared elk.
- indivSocConnectedness: Scaled eigenvector centrality of the focal individual, based on proximity-based social networks of all collared elk for the 14 days prior to the step.
- indivSocFamiliarity: Median simple ratio index of dyadic association between focal individual and its group members, based on proximity-based social networks of all collared elk for the 14 days prior to the step.
- indivSocStability: Median number of hours since the fusion of dyads between the focal individual and its group members.


[Social phenotypes of groups]

- groupDominance: Mean elo-score of most dominant individual in the group (i.e., maximum elo-score of group members), based on observations of agonistic interactions between pairs of collared elk.
- groupSocConnectedness: Median scaled eigenvector centrality of all group members, based on proximity-based social networks of all collared elk for the 14 days prior to the step.
- groupSocFamiliarity: Median simple ratio index of all dyadic associations between group members, based on proximity-based social networks of all collared elk for the 14 days prior to the step.
- groupSocStability: Median number of hours since the fusion of all dyads of group members.


# Code

## 0) Load data and set working directory

```{r Libraries}

# Libraries 
library(ggplot2)
library(tidyverse)
library(tidyr)
library(glmmTMB)

```

```{r setup, include=FALSE}

# Set working directory
knitr::opts_knit$set(root.dir = "G:/SET WORKING DIRECTORY HERE/")

```

```{r Load travel steps ready for models}

allData <- read.csv("./PoulinEtAl2023-hourlyTravelSteps_wSocialPhenotypes.csv")

```


## 1) Traffic volume

Run a model with only traffic volume and season as predictors of highway crossings.

```{r}

# Run model
m1 <- glmmTMB(hwyCros ~ trafficVol +
                season +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m1)

# Get R^2
MuMIn::r.squaredGLMM(m1) 

```


## 2) Group size

Run a model with group size, traffic volume and season as predictors of highway crossings.

```{r}

# Run models
m2 <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m2)

# Get R^2
MuMIn::r.squaredGLMM(m2) 

```


## 3) Add social phenotypes to models of highway crossings

Run models with social phenotypes of individuals or groups, with group size, traffic volume and season as predictors of highway crossings.


### Dominance

```{r Social phenotype of individuals}

# Run models
m3_indivDom <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                indivDominance +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_indivDom)

# Get R^2
MuMIn::r.squaredGLMM(m3_indivDom)

```

```{r Social phenotype of groups}

# Run models
m3_groupDom <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                groupDominance +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_groupDom)

# Get R^2
MuMIn::r.squaredGLMM(m3_groupDom)

```

```{r Compare models with AIC}

AIC(m3_indivDom, m3_groupDom)

```


### Social connectedness

```{r Social phenotype of individuals}

# Run models
m3_indivSocCon <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                indivSocConnectedness +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_indivSocCon)

# Get R^2
MuMIn::r.squaredGLMM(m3_indivSocCon)

```

```{r Social phenotype of groups}

# Run models
m3_groupSocCon <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                groupSocConnectedness +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_groupSocCon)

# Get R^2
MuMIn::r.squaredGLMM(m3_groupSocCon)

```

```{r Compare models with AIC}

AIC(m3_indivSocCon, m3_groupSocCon)

```


### Social familiarity

```{r Social phenotype of individuals}

# Run models
m3_indivSocFam <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                indivSocFamiliarity +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_indivSocFam)

# Get R^2
MuMIn::r.squaredGLMM(m3_indivSocFam)

```

```{r Social phenotype of groups}

# Run models
m3_groupSocFam <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                groupSocFamiliarity +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_groupSocFam)

# Get R^2
MuMIn::r.squaredGLMM(m3_groupSocFam) # 0.30267100

```

```{r Compare models with AIC}

AIC(m3_indivSocFam, m3_groupSocFam)

```


### Social stability

```{r Social phenotype of individuals}

# Run models
m3_indivSocStab <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                indivSocStability+
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_indivSocStab)

# Get R^2
MuMIn::r.squaredGLMM(m3_indivSocStab)

```

```{r Social phenotype of groups}

# Run models
m3_groupSocStab <- glmmTMB(hwyCros ~ trafficVol +
                season +
                groupSize +
                groupSocStability +
                (1|elkID_winterSeason),
              data = allData,
              family = "binomial")

# Look at summary
summary(m3_groupSocStab)

# Get R^2
MuMIn::r.squaredGLMM(m3_groupSocStab)

```

```{r Compare models with AIC}

AIC(m3_indivSocStab, m3_groupSocStab)

```


# 4) Interaction between social phenotypes and traffic volume

Run models with social phenotypes of individuals or groups in interaction with traffic volume, as well as group size, and season as predictors of highway crossings.


### Group size

```{r}

# Run models
m4_groupSize <- glmmTMB(hwyCros ~ trafficVol100*groupSize +
                season +
                (1|elkID_winterSeason),
                data = allData,
                family = "binomial")

# Look at summary
summary(m4_groupSize)

# Get R^2
MuMIn::r.squaredGLMM(m4_groupSize)

```


### Dominance

```{r}

# Run models
m4_dom <- glmmTMB(hwyCros ~ trafficVol100*groupDominance +
                    groupSize +
                    (1|elkID_winterSeason),
                    data = allData,
                    family = "binomial")

# Look at summary
summary(m4_dom) 

# Get R^2
MuMIn::r.squaredGLMM(m4_dom)

```


### Social connectedness

```{r}

# Run models
m4_socCon <- glmmTMB(hwyCros ~ trafficVol100*groupSocConnectedness +
                    groupSize +
                    (1|elkID_winterSeason),
                    data = allData,
                    family = "binomial")

# Look at summary
summary(m4_socCon)

# Get R^2
MuMIn::r.squaredGLMM(m4_socCon)

```


### Social familiarity

```{r}

# Run models
m4_socFam <- glmmTMB(hwyCros ~ trafficVol100*groupSocFamiliarity +
                    groupSize +
                    (1|elkID_winterSeason),
                    data = allData,
                    family = "binomial")

# Look at summary
summary(m4_socFam) 

# Get R^2
MuMIn::r.squaredGLMM(m4_socFam) 

```


### Social stability

```{r}

# Run models
m4_socStab <- glmmTMB(hwyCros ~ trafficVol100*indivSocStability +
                    groupSize +
                    (1|elkID_winterSeason),
                    data = allData,
                    family = "binomial")

# Look at summary
summary(m4_socStab)

# Get R^2
MuMIn::r.squaredGLMM(m4_socStab)

```

